---
title: "STAA57 Assignment"
author: "Group"
date: "2024-03-28"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
# Read the CSV file located at the specified path into a data frame named 'd'
d =read.csv ("/Users/mark/Desktop/course/staa57/R_CSV_DATA_ANALYZE/g0811_05.csv") 
# Load the 'tidyverse' package, which includes a collection of R packages for data manipulation and visualization
library(tidyverse)
# Load the 'ggplot2' package, a popular R package for creating graphics and plots
library(ggplot2)
# Load the 'knitr' package, which provides tools for dynamic report generation in R
library(knitr)
# Load the 'rpart' package, which provides functions for recursive partitioning and regression trees
library(rpart)
```


## 1.Description
```{r}
# Display the column names of the data frame 'd'
names(d)
```
(describe variables)

## 2.The Background of the Data
https://data.ontario.ca/dataset/reason-for-part-time-work-by-immigration-status

## 3.What is the over all research question？
1. Do more people work part-time as the years go by?
2. Are there any significant differences in part-time population growth between Canada and Ontario


## 4. Tables
# 4.1 Total population of part-time job over Canada

```{r}
# Create a new data frame 'd_table1' by filtering the data frame 'd' to select rows where:
# - Reason is "Part-time employment, all reasons"
# - Immig. is "Total"
# - Age.group is "15+"
# Then, select only the columns 'Year', 'Geography', 'Both.sexes', 'Men', and 'Women' from the filtered data frame.
# Arrange the resulting data frame by the 'Geography' column.
d_table1=d %>%
  filter(Reason=="Part-time employment, all reasons" &Immig.=="Total" &Age.group=="15+") %>% select(Year,Geography,Both.sexes,Men,Women) %>% arrange(Geography)
# Display the data in 'd_table1' as a nicely formatted table using the 'kable' function
kable(d_table1)
```

# 4.2（test reason=	Caring for children 的比例 ，男女的ratio是否一样（4.3 table）） test ratio=0.5
```{r}
# Filter the data frame 'd' to select rows where 'Geography' is "Canada"
d_ratio= d %>% filter(Geography=="Canada") 

# Further filter the 'd_ratio' data frame to select rows where:
# - 'Reason' is "Caring for children"
# - 'Age.group' is "15+"
# - 'Immig.' is "Total"
# Then, calculate the ratio of 'Men' to 'Both.sexes' and create a new column named 'ratio'.
# Finally, select the 'Year' and 'ratio' columns.
d_ratio=d_ratio %>%
  filter(Reason=="  Caring for children" & Age.group=="15+" & Immig.=="Total" ) %>% 
  mutate(ratio=Men/Both.sexes) %>% select(Year,ratio) 

# Display the data in 'd_ratio' as a nicely formatted table using the 'kable' function
kable(d_ratio)

```


## 5. Graphs
# 5.1 Graph of population of par-time job men over years
```{r}
# Create a ggplot object and add a line plot using the 'geom_line' function.
# Map the 'Year' variable to the x-axis, 'Men' variable to the y-axis, and 'Geography' variable to the color aesthetic.
# Use data from the 'd_table1' data frame.
# Label the y-axis as "population of part-time job men"
# Set the title of the plot as "Graph of population of par-time job men over years"
ggplot()+geom_line(aes(x=Year,y=Men,col=Geography),data=d_table1)+
  ylab("population of part-time job men")+
  ggtitle("Graph of population of par-time job men over years")

```


# 5.2 Graph of population of part-time job for women over year
```{r}
# Create a ggplot object and add a scatter plot using the 'geom_point' function.
# Map the 'Year' variable to the x-axis and the 'ratio' variable to the y-axis, using data from the 'd_ratio' data frame.
# Label the y-axis as "population of part-time job for women"
# Add a linear regression line to the plot using the 'geom_smooth' function
# Set the title of the plot
ggplot()+geom_point(aes(x=Year,y=ratio),data=d_ratio)+
  ylab("population of part-time job for women")+
  geom_smooth(method="lm")+  #why no line
  ggtitle("Graph of population of part-time job for women over years")
```



## 6. Confidence interval and test of hypothesis
test reason=caring for children 的比例 ，男女的proportion是否一样（4.3 table）
H0: 男的比率是0.5 也就是男女比例都一样
HA：more women ie ratio<0.5
```{r}
# Perform a one-sample t-test on the 'ratio' variable from the 'd_ratio' data frame.
# The null hypothesis (H0) is that the population mean of the 'ratio' variable is equal to 0.5.
# The alternative hypothesis (H1) is that the population mean of the 'ratio' variable is less than 0.5.
# Use a confidence level of 95%.
t.test(d_ratio %>% select(ratio),alternative="less",mu=0.5,conf.level=0.95)
```
reject  H0 more women 因为caring for children而选择parttime

## 7. Bootstrapping

# 7.1 Bootstrapping Analysis for Slope Coefficient Estimation in Linear Regression Model
The code below is to perform bootstrapping to estimate the variability of the slope coefficient in a linear regression model where 'Both.sexes' is the response variable and 'Year' is the predictor variable. The code achieves this by repeatedly sampling with replacement from the dataset ('d_table1'), fitting linear models to the sampled data, extracting the slope coefficient from each model, and finally calculating the 2.5th and 97.5th percentiles of the slope coefficient estimates obtained from bootstrapping.
```{r}
# Define a bootstrapping function named 'boot_function'.
# Within the function:
# - Sample with replacement from the 'd_table1' data frame and store it in 'boot.d'.
# - Fit a linear model ('lm') with 'Both.sexes' as the response variable and 'Year' as the predictor variable using the sampled data.
# - Extract the slope coefficient ('s') from the fitted model.
# - Return the slope coefficient.
boot_function=function(){
  boot.d = d_table1 %>% sample_n(nrow(d_table1), replace=T)
  m2 = lm(Both.sexes~Year, data=boot.d)
  s = coef(m2)[2]
  return(s) 
}
# Perform bootstrapping by calling 'boot_function' 1000 times and store the output in 'output'.
output = replicate(1000, boot_function()) 

# Calculate the 2.5th and 97.5th percentiles of the 'output' vector.
quantile(output, c(0.025,0.975))

```
**From the above consequence:**

    ·The 2.5th percentile of the bootstrapped slope coefficients is approximately -73.7383.
    ·The 97.5th percentile of the bootstrapped slope coefficients is approximately 124.1169.
  The positive relationship between 'Year' and 'Both.sexes' suggests a potential long-term trend in the variable 'Both.sexes' over the years covered in the dataset. This trend could indicate societal changes, cultural shifts, or policy interventions that have led to an increase in the value of 'Both.sexes' over time.
  
  In the meanwhile, the increasing values of 'Both.sexes' may reflect progress towards gender equality or a blurring of traditional gender roles over time. This trend could be indicative of greater acceptance and recognition of diverse gender identities and expressions in society.
  
  By the way, this ouput may also contribute to raising public awareness about the evolving nature of gender identities and expressions. It underscores the importance of fostering inclusive environments and promoting acceptance and understanding of diverse gender experiences.

# 7.2 Bootstrapping Analysis for Mean Estimation
The code below is to conduct bootstrapping to estimate the mean of the 'ratio' variable in the 'd_ratio' dataset. The function 'boot_function2' is defined to repeatedly sample with replacement from the 'ratio' variable, calculate the mean of the sampled data, and return it. The code then performs bootstrapping by calling 'boot_function2' 1000 times and calculates the 5th percentile of the resulting output vector.
```{r}
# Define a bootstrapping function named 'boot_function2'.
# Within the function:
# - Select the 'ratio' variable from the 'd_ratio' data frame and store it in 'obs.sample'.
# - Sample with replacement from 'obs.sample' and store it in 'boot.d'.
# - Calculate the mean of the sampled data and return it.
boot_function2=function(){
  obs.sample=d_ratio %>% select(ratio)
  boot.d = sample_n(obs.sample,size=nrow(d_ratio), replace=T)
  return(mean(boot.d[,1])) 
}

# Perform bootstrapping by calling 'boot_function2' 1000 times and store the output in 'output'.
output = replicate(1000, boot_function2())

# Calculate the 5th percentile of the 'output' vector.
quantile(output,probs=c(0.05))

```
**From the above consequence:**

    The output indicates that the 5th percentile of the bootstrapped means of the 'ratio' variable is approximately 0.045. This suggests that there is a 5% probability that the true mean of the 'ratio' variable in the population is lower than or equal to 0.045, based on the bootstrapped samples.
  
    Therefore, the conclusion is that the estimated mean of the 'ratio' variable in the dataset 'd_ratio' is likely to be greater than 0.045, with 95% confidence. This information can be valuable for understanding the central tendency of the 'ratio' variable and making inferences about the population mean based on the sampled data.



## 8 Regression analysis
# 8.1 random forest
```{r}
# Filter the data frame 'd' to exclude rows where 'Reason' is "Part-time employment, all reasons".
d1=d %>% filter(Reason!="Part-time employment, all reasons")
# Assign a binary value to the 'reason' variable based on 'Reason' values:
# - 0 if the reason does not involve looking for full-time work in the last month
# - 1 if the reason involves looking for full-time work in the last month
d1=d1%>% mutate(reason=case_when(Reason=="  Own illness"~0,
                               Reason=="  Caring for children"~0,
                               Reason=="  Other personal or family responsibilities"~0,
                               Reason=="  Going to school"~0,
                               Reason=="  Personal preference"~0,
                               Reason=="  Other voluntary"~0,
                               Reason=="  Business conditions, did not look for full-time work in last month"~0,
                               Reason=="  Could not find full-time work, did not look for full-time work in last month"~0,
                               Reason=="  Business conditions, looked for full-time work in last month"~1,
                               Reason=="  Could not find full-time work, looked for full-time work in last month"~1))
# Load the 'randomForest' package for building random forest models
library(randomForest)
# Build a random forest model ('rforest.m') to predict the 'reason' variable based on other variables.
rforest.m = randomForest(reason ~ Year+Geography+Immig.+Age.group++Both.sexes+Men+Women,data=d1,ntree=500,importance=TRUE)
# Visualize variable importance using the 'varImpPlot' function
varImpPlot(rforest.m)
# Make predictions using the random forest model on the same data ('d1') and calculate the mean squared error
d_randomforest=d1%>%mutate(rforest_pred=predict(rforest.m,new_data=d1))

mean((d_randomforest$reason-d_randomforest$rforest_pred)^2)

```

# 8.2 Linear regression
```{r}
# Fit a linear regression model ('m2') with 'Both.sexes' as the response variable and 'Year' as the predictor variable using data from 'd_table1'.
m2=lm(Both.sexes~Year,data=d_table1)
# Display a summary of the linear regression model
summary(m2)
```
Here are some explanations about the component:


 (解释pvalue， intercept，slope)
 for d_table1, 0.9248 of the population can be explained by the linear regression model.
 

# 9 Summary of Research
Based on the information presented in the report, the following are the key findings:


